Parallel air palletization and tour planning for simultaneous pickup and delivery

### 2023-03-27 latex/comp_or.tex
Paper CompOR update

The other parameter $volume$\/ is a value greater than 1, which corresponds, at each node $\pi_k$, to the ratio between the sum of the volumes of the items ($\sum_{j=1}^{n^{\pi_k}} t^{\pi_k}_j.v$) and the load capacity of the pallets ($\sum_{i=1}^{m} p_i.V$). This parameter is passed to $ItemsGeneration$ (line \ref{main:items}), responsible for creating the items to be shipped, which will be presented in the next section.

$method$\/ corresponds to one of the heuristics that we will present in subsection \ref{methods}. The loop of lines \ref{main:loop1}-\ref{main:loop2} goes through all permutations $\pi$, where the node-by-node resolutions are performed by $SolveTour$, whose result is stored in $f_{\pi}$. The best outcome among all $K!$\/ tours will be the answer for $scenario$, $volume$\/ and $method$\/ (line \ref{main:f}).

@@

In addition to the set of nodes, pallets, costs and items, $SolveTour$, described in Algorithm \ref{alg:tour}, receives the parameter $method$, which corresponds to a heuristic for solving the node-by-node problems, and the parameter $\pi$, which is a permutation that defines the order of visits in this tour.

As we mentioned in the previous section, all tours start and end at $\pi_0$\/ (lines \ref{tour:pi1}-\ref{tour:pi2}).
After initializing the score and cost values (lines \ref{tour:score}-\ref{tour:cost}), there is a loop for the $K+1$\/ flights (lines \ref{tour:loop1}-\ref{tour:loop2}). Initially we set pallets destination as $-1$\/ (line \ref{tour:-1}). When the aircraft is at node $\pi_0$, the initial graph $G_1$\/ is empty because it has no consolidated items \ref{tour:g11}. Otherwise, the set $L_k$\/ of remaining nodes when departing from $\pi_k$ is updated (line \ref{tour:lk1}), and $UpdateConsolidated$\/ (line \ref{tour:dest}) returns the set of consolidated items that have not yet reached their destinations and remain on board, rearranging them on the pallets to minimize CG deviation. This allocation is stored in graph $G_1$\/ (line \ref{tour:g12}).

@@

		\State $\pi[0] \gets 0$ \label{tour:pi1}
		\State $\pi(K+1) \gets 0$ \label{tour:pi2}
		\State $score \gets 0$ \label{tour:score}
		\State $cost \gets 0$ \label{tour:cost}
		\For {$k\gets0$ to $K$} \label{tour:loop1}		
			\For{$i \gets 1$ to $m$}
				\State $p_i.to[k] \gets -1$ \label{tour:-1} \Comment{reset this pallet destination}
			\EndFor	
			\If {$k = 0$}
				\State $L_0 \gets L$
				\State Let $G_1(M \cup N_0, \varnothing)$ \label{tour:g11}
			\Else
				\State $L_k \gets L_k - \pi(k)$  \label{tour:lk1}			
				\State $Q^{\pi_k}, E_Q^{\pi_k}, M^{\pi_k} \gets UpdateConsolidated(\pi(k))$ \label{tour:dest}			
				\State Let $G_1(M^{\pi_k} \cup N^{\pi_k} \cup Q^{\pi_k}, E_Q^{\pi_k})$ \label{tour:g12}
			\EndIf  \label{tour:lk2}	
			\State $M^{\pi_k} \gets SetPalletsDestination( \pi(k) )$ \label{tour:dest2}		
			\State $G_2 \gets SolveNode(method, \pi(k), G_1, timeLim)$ \label{tour:node}

{\color{purple} \State	$G_3 \gets 3Dpacking(G_2)$  \label{tour:3dpacking} }

@@

In the context of this work, we know that $m>K$, once the aircraft has $7$\/ or $18$\/ pallets and $K\leq 6$, allowing there to be at least one pallet for each node to be visited. $SetPalletsDestination$\/ (line \ref{tour:dest2}) presets the destination of each pallet based on the volume demands caused by the items to be embarked in the current node, without changing the pallets destination with consolidated items.

@@

$SetPalletsDestination$ sets pallets destination not yet defined. The total volume of items and the number of pallets with consolidated items destined for each node. The destinations of each pallet are defined proportionally to the volume of items, regarding the pallets with consolidated items.

{\color{purple}
In line \ref{tour:3dpacking}, $3Dpacking(G)$ packs the items previously assigned to a pallet, which is feasible in terms of volume, weight, and torque but still not guaranteed to fit into the pallet. In this procedure, some unfit items are excluded from the solution.
}

@@

$UpdateConsolidated$, $MinCGDeviation$, $SetPalletsDestination$, and $ScoreAndDeviation$ are thoroughly described by \cite{MesquitaSanches2023}.

@@

According to \cite[p.226]{manfrin2006}, there are five topologies for multiprocessing interchange of information: {\it fully-connected}, where the master process broadcasts to all remaining child processes; {\it replace-worst}, where the best-so-far solution process broadcasts only to the current worst solution process; {\it hypercube}, where processes are connected as a hypercube, and a vertex process broadcasts only to the connected vertices; {\it ring}, in which one process only sends a message to the next process connected to it; and {\it parallel independent runs (PIR)}, in which there are no communication costs and the best solution is chosen among all processes. They concluded that the PIR topology presented, in average, the smallest distance from the optimal solutions.

In this work, we had to select the {\it fully-connected} topology because the list of items to embark and the aircraft torque must be accessible for reading and writing by all child processes.

@@

In the first and third phases, we considered each pallet as a parallel process, taking advantage of nowadays multi-core computers.

@@

All phases being executed in a multiprocessing run-time, accessing concurrently, reading and writing the list of items and the current value of the CG deviation.

@@

The main logic of the 3D packing procedure is based on an approximate algorithm like follows:

\begin{itemize}
	\item[1] From a list of items assigned to a pallet, items are sorted from the biggest to the smallest and placed in such an order on a pallet. Each item has 1 to 6 orientations to choose from when it can be placed. The orientation procedure can select the best direction type among possible orientations.
	\item[2] A pivot point is used to determine item's position. The pivot is an (x, y, z) coordinate which represents a point in the pallet at which an attempt to pack an item will be made.
	\item[3] The back lower left corner of the item will be placed at the pivot. If the item cannot be packed at the pivot position then it is rotated until it can be packed at the pivot point or until we have tried all 6 possible rotation types.
	\item[4] If, after rotating it, the item still cannot be packed at the pivot point, then we move on to packing another item and add the unpacked item to a list of items that will be packed after an attempt to pack the remaining items is made.
	\item[5] The first pivot in the empty pallet is always (0, 0, 0). When one item can be placed into multiple optimal pivot point, the placement selection module can help make a choice.
\end{itemize}

@@

Let $p_i.Dh,\ p_i.Wh,\ p_i.H$ be parameters indicating the depth, width, and height of the pallet. We included letter $h$ at the end of parameters $p_i.Dh$ and $p_i.Wh$ to turn them different from $p_i.D$ (distance from CG) and $p_i.W$ (weight limit).

@@

\begin{equation} \label{eq:3dpacking1}
	\mbox{minimize} \Big [ p_i.Dh \times p_i.Wh \times p_i.H - \sum_{j \gets 1}^{n^{\ i}} X_{ij} \times A_{j} \times t_j.d \times t_j.wth \times t_j.h \Big ]
\end{equation}

\begin{equation} \label{eq:3dpacking2}
	\sum_{j \gets 1}^{n^{\ i}} A_{j} \times t_j.d  \leq X_{ij} \times p_i.D\ i \in \{1,\ 2,\ 3,\ ...,\ m\}
\end{equation}

@@

\begin{equation} \label{eq:3dpacking5}
t_j.o^1 + t_j.o^2 + t_j.o^3 + t_j.o^4 + t_j.o^5 + t_j.o^6 = 1
\end{equation}

Equation \ref{eq:3dpacking1} minimizes the volume slack; equations \ref{eq:3dpacking2}, \ref{eq:3dpacking3}, and \ref{eq:3dpacking4} constrains the packing to the three dimensions; and \ref{eq:3dpacking5} guarantees that only one orientation is chosen for the item.

@@

\begin{algorithm}[H]
	\caption{Parallel 3D packing procedure}  \label{alg:3dpacking}

@@

\subsection{Minimization of the distance of a next node destined pallet to the ramp door }

@@

@@

\subsection{Results obtained}

@@

\begin{table}[H]
	\centering
	\caption{Volume surpluses, methods and scenarios results {\bf before} 3D packing}  \label{tab:results}
	\footnotesize
	%\scriptsize
	\begin{tabular}{crcccccccc}
		\toprule
		{\bf Volume}     &  {\bf Resolution}   &                              &\multicolumn{6}{c}{\bf Scenarios}                                                   &{\bf Normalized}\\
		\cmidrule{4-9}		
		{\it surplus}    & {\it method}        & {\bf Results}                &{\bf 1}    &{\bf 2}    &{\bf 3}    &{\bf 4}    &{\bf 5}    &{\bf 6}    &{\bf Speedup} \\
		\toprule

		%\specialcell{$Volume$\\$surplus$}

		\multirow{6}{*}{$1.2$}&$Relaxed$                 & $f$         &{\it 12.23}&{\it 8.67} &{\it 12.67}&{\it 15.36}&{\it 16.29}&{\it 16.20}&  \specialcell{{\bf 1.00}\\-} \\%  8142
		&                      $MIP$                     & {\it run time (s)} &     3     &     13    &     20    &     78   &      198  &     1103  &  1.00 \\% 439							
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $Shims$}                       & $f$                &     10.34 &     8.50  &     12.42 &     15.06 &     15.97 &     15.88 &  0.96 \\% 7817
		&                                                & {\it run time (s)} &     < 1   &     3     &     5     &     20    &     58    &     352   &  1.00 \\% 439			
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $mpShims$}                     & $f$                &   12.96   &   7.53    &  12.28    &   13.20   &  13.71    &   13.44   &  0.90\\% 7278
		&                                                & {\it run time (s)} &     < 1   &     2     &     2     &     8    &     22    &     120   &  {\bf 2.80}\\%

		\midrule

		\multirow{6}{*}{$1.5$}&\specialcell{$Relaxed$\\$MIP solver$}& $f$         &{\it 17.01}&{\it 12.45}&{\it 18.02}&{\it 22.03}&{\it 23.09}&{\it 22.35}&  \specialcell{{\bf 1.00}\\-} \\%11495

        \cmidrule{2-10}		                       
        &\multirow{2}{*}{ $Shims$}                       & $f$                &     14.68 &     12.21 &     17.67 &  21.01    &  22.64    &   21.91   &  0.96  \\%  11012
        &                                                & {\it run time (s)} &     < 1   &     2     &     3     &     12    &     33    &   209     &  1.00 \\% 260

        \cmidrule{2-10}		                       
        &\multirow{2}{*}{ $mpShims$}                     & $f$                &    16.12  &     11.81 &     17.48 &  17.66    &   19.81   &  17.61    &  0.87 \\% 9966
        &                                                & {\it run time (s)} &     < 1   &     1     &     2     &     7     &   20      &  120      &  {\bf 1.70}\\% 151

		\midrule
		\multirow{6}{*}{$2.0$}&\specialcell{$Relaxed$\\$MIP solver$}& $f$         &{\it 24.56}&{\it 17.47}&{\it 25.19}&{\it 27.94}&{\it 29.90}&{\it 27.60 }&  \specialcell{{\bf 1.00}\\-} \\%15266

		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $Shims$}                       & $f$                &    21.31  &  17.35    &   24.24   &   27.39    &  29.31     &  27.06     &   0.96  \\%  14666
		&                                                & {\it run time (s)} &     < 1   &     2     &     4     &     14     &     35     &   220     &  1.00 \\% 260

		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $mpShims$}                     & $f$                &    24.33  &   17.26   &   24.70   &  24.81     &   27.71    &  24.93        &  0.94  \\% 14374
		&                                                & {\it run time (s)} &     < 1   &     2     &     3     &     9      &   26       &  151      &  {\bf 1.70}\\% 151

		\bottomrule

	\end{tabular}
	%\textsuperscript{asdf sd .}	
	\normalsize
\end{table}

@@

\begin{table}[H]
	\centering
	\caption{Volume surpluses, methods and scenarios results {\bf after} 3D packing}  \label{tab:results2}
	\footnotesize

\color{blue}{

	%\scriptsize
	\begin{tabular}{crcccccccc}
		\toprule
		{\bf Volume}     &  {\bf Resolution}   &                              &\multicolumn{6}{c}{\bf Scenarios}                                                   &{\bf Normalized}\\
		\cmidrule{4-9}		
		{\it surplus}    & {\it method}        & {\bf Results}                &{\bf 1}    &{\bf 2}    &{\bf 3}    &{\bf 4}    &{\bf 5}    &{\bf 6}    &{\bf Speedup} \\
		\toprule

		%\specialcell{$Volume$\\$surplus$}

		\multirow{6}{*}{$1.2$}&\specialcell{$Relaxed$\\$MIP solver$}& $f$            &{\it 11.47}&{\it 7.24} &{\it 10.78}&{\it 12.06}&{\it 12.97}&{\it 12.44}&  \specialcell{{\bf 1.00}\\-} \\%  8142
		&                                                & {\it run time (s)} &     6     &     25    &     41    &     158   &      401  &     2161  &  1.00 \\% 439					
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $Shims$}                       & $f$                &    9.05   &   5.11    &   9.75    &   10.96   &   12.86   &   13.25   &  0.96 \\% 7817
		&                                                & {\it run time (s)} &     16    &     68    &     94    &    323    &    719    &    3549   &  1.00 \\% 439			
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $mpShims$}                     & $f$                &    12.14  &   7.29    &   10.36   &   12.63   &   14.04   &   13.92   &  0.90\\% 7278
		&                                                & {\it run time (s)} &     9     &     36    &     54    &     192   &     427   &    2118   &  {\bf 2.80}\\%

		\midrule

		\multirow{6}{*}{$1.5$}&\specialcell{$Relaxed$\\$MIP solver$}& $f$     &{\it 15.83}&{\it 10.08}&{\it 15.14}&{\it 16.87}&{\it }&{\it }&  \specialcell{{\bf 1.00}\\-} \\%11495  
		&                                                & {\it run time (s)} &    10     &     43    &     72    &     262   &     58    &     352   &  1.00 \\% 439					
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $Shims$}                       & $f$                &   11.55   &   7.61   &   12.28   &   15.00   &   17.55   &   17.57   &  0.96  \\%  11012
		&                                                & {\it run time (s)} &    25     &    107   &     156   &    498    &    1217   &   5743     &  1.00 \\% 260

		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $mpShims$}                     & $f$                &   17.05  &   10.01   &   15.01  &   16.75   &   20.09   &      &  0.87 \\% 9966
		&                                                & {\it run time (s)} &    18    &    64     &     101  &    351    &   828     &  120      &  {\bf 1.70}\\% 151

		\midrule
		\multirow{6}{*}{$2.0$}&\specialcell{$Relaxed$\\$MIP solver$}& $f$            &{\it }&{\it }&{\it }&{\it }&{\it }&{\it  }&  \specialcell{{\bf 1.00}\\-} \\%15266  
		&                                                & {\it run time (s)} &     < 1   &     3     &     5     &     20    &     58    &     352   &  1.00 \\% 439					
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $Shims$}                       & $f$                &      &      &      &       &       &       &   0.96  \\%  14666
		&                                                & {\it run time (s)} &     < 1   &     2     &     4     &     14     &     35     &   220     &  1.00 \\% 260

		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $mpShims$}                     & $f$                &      &      &      &       &       &          &  0.94  \\% 14374
		&                                                & {\it run time (s)} &     < 1   &     2     &     3     &     9      &   26       &  151      &  {\bf 1.70}\\% 151

		\bottomrule

	\end{tabular}
}
	%\textsuperscript{asdf sd .}	
	\normalsize
\end{table}

@@

We confirmed the hypothesis raised in \ref{hypotheses} and in line \ref{mpShims:minCGdev} of Algorithm \ref{alg:mpShims} that, in sequential mode, where one pallet is built at a time, torque constraints could be more restrictive than volume or weight constraints, which could stop the palletization process earlier, yielding a solution that may be not as good as it could be. This integer programming procedure inclusion really improved all {\it Shims} results.

@@

\subsection{Results of the 3D packing}

%\begin{table}[!h]
%	\begin{minipage}{0.5\linewidth}
%		left
%	\end{minipage}\hfill % these two lines must be close to each other
%	\begin{minipage}{0.5\linewidth}
%		right		
%	\end{minipage}
%\end{table}

\vspace{2.0mm}
\begin{table}[!ht]
\centering
\caption{Parallel 3D packing time increment}  \label{tab:3d}
\scriptsize
\begin{tabular}{c S[table-format=-1.1] S[table-format=-1.1] S[table-format=-1.1]}
\toprule
&   \multicolumn{3}{c}{\bf Average tour time - 3D packing}  \\
\cmidrule{2-4}
& {\bf without} & {\bf with} & {\bf increment} \\
{\bf Scenario} & {\bf (s)}    & {\bf (s)}   & {\bf (times)} \\		
\midrule
$1$ &   1 &   15 & 15 \\ % 1	 9   , 3	 6, 1	 25	
$2$ &   5 &   76 & 15 \\ % 2	 36  , 13	 25
$3$ &   8 &   97 & 12 \\ % 3	 54  , 20	 41	
$4$ &  35 &  330 &  9 \\ % 12	 192 , 78	 158
$5$ &  98 &  740 &  8 \\ % 34	 427 , 198	 401
$6$ & 591 & 3709 &  6 \\ % 202	 2118, 1103	 2161                                                       
\bottomrule
\end{tabular}
\end{table}

According to Table \ref{tab:3d}, the time increase with the 3D packing inclusion was approximately 10 times. Considering the worst (6) scenario, about 62 min to solve a complete tour, against 10 min without the 3D packing procedure.
i.e., with the 3D packing procedure running in multiprocessing mode, this complete solution method for {\it ACLP+RPDP} is still viable for operational use.

\subsection{Results of the minimization of the pallet handling distances in the cargo bay}
\label{improvement}

>>>>

Argument $method$ corresponds to one of the heuristics that we will present in subsection \ref{methods}.

The loop of lines \ref{main:loop1}-\ref{main:loop2} goes through all permutations $\pi$, where the node-by-node resolutions are performed by $SolveTour()$, whose result is stored in $f_{\pi}$.

The best outcome among all $K!$ tours will be the answer for $scenario$, $volume$ and $method$ (line \ref{main:f}).

@@

In addition to the set of nodes, pallets, costs and items, $SolveTour$, described in Algorithm \ref{alg:tour}, receives the parameter $method$, which corresponds to a heuristic for solving the node-by-node sub-problems, and the parameter $\pi$, which is a permutation that defines the order of visits in this tour.

As we mentioned in the previous section, all tours start and end at $\pi_0$\/ (lines \ref{tour:pi1}-\ref{tour:pi2}).
After initializing the score and cost values (lines \ref{tour:score}-\ref{tour:cost}), there is a loop for the $K+1$\/ tours (lines \ref{tour:loop1}-\ref{tour:loop2}). Initially we set pallets destination as $-1$\/ (line \ref{tour:-1}). When the aircraft is at node $\pi_0$, the initial graph $G_1$\/ is empty because it has no consolidated items \ref{tour:g11}. Otherwise, the set $L^{\pi_k}$\/ of remaining nodes when departing from $\pi_k$ is updated (line \ref{tour:lk1}), and $UpdateConsolidated$\/ (line \ref{tour:dest}) returns the set of consolidated items that have not yet reached their destinations and remain on board, rearranging them on the pallets to minimize CG deviation. This allocation is stored in graph $G_1$\/ (line \ref{tour:g12}).

@@

		\State $\pi[0] \gets 0$ \label{tour:pi1}
		\State $\pi(K+1) \gets 0$ \label{tour:pi2}
		\State $score \gets 0$ \label{tour:score}
		\State $cost \gets 0$ \label{tour:cost}
		\For {$k\gets0$ to $K$} \label{tour:loop1}		
			\For{$i \gets 1$ to $m$}
				\State $p_i.to[k] \gets -1$ \label{tour:-1} \Comment{reset this pallet destination}
			\EndFor	
			\If {$k = 0$}
				\State $L_0 \gets L$
				\State Let $G_1(M \cup N_0, \varnothing)$ \label{tour:g11}
			\Else
				\State $L_k \gets L_k - \pi(k)$  \label{tour:lk1}			
				\State $Q^{\pi_k}, E_Q^{\pi_k}, M^{\pi_k} \gets UpdateConsolidated(\pi(k))$ \label{tour:dest}			
				\State Let $G_1(M^{\pi_k} \cup N^{\pi_k} \cup Q^{\pi_k}, E_Q^{\pi_k})$ \label{tour:g12}
			\EndIf  \label{tour:lk2}	
			\State $M^{\pi_k} \gets SetPalletsDestinations( \pi(k) )$ \label{tour:dest2}		
			\State $G_2 \gets SolveNode(method, \pi(k), G_1, timeLim)$ \label{tour:node}

{\color{purple} \State	$G_3 \gets mp3Dpacking(G_2)$  \label{tour:3dpacking} }

@@

In the context of this work, we know that $m>K$, once the aircraft has $7$\/ or $18$\/ pallets and $K\leq 6$, allowing there to be at least one pallet for each node to be visited. $SetPalletsDestinations$\/ (line \ref{tour:dest2}) presets the destination of each pallet based on the volume demands caused by the items to be embarked in the current node, without changing the pallets destination with consolidated items.

@@

$SetPalletsDestinations$ sets pallets destination not yet defined. The total volume of items and the number of pallets with consolidated items destined for each node. The destinations of each pallet are defined proportionally to the volume of items, regarding the pallets with consolidated items.

{\color{purple}
In line \ref{tour:3dpacking}, $mp3Dpacking(G)$ packs the items previously assigned to a pallet, which is feasible in terms of volume, weight, and torque but still not guaranteed to fit into the pallet. In this procedure, some unfit items are excluded from the solution.
}

@@

$UpdateConsolidated$, $MinCGDeviation$, $SetPalletsDestinations$, and $ScoreAndDeviation$ are thoroughly described by \cite{MesquitaSanches2023}.

@@

According to \cite[p.226]{manfrin2006}, there are five topologies for multiprocessing interchange of information: {\it fully-connected}, where the master process broadcasts to all remaining child processes; {\it replace-worst}, where the best-so-far solution process broadcasts only to the current worst solution process; {\it hypercube}, where processes are connected as a hypercube, and a vertex process broadcasts only to the connected vertices; {\it ring}, in which one process only sends a message to the next process connected to it; and {\it parallel independent runs}, in which there are no communication costs and the best solution is chosen among all processes.

In this work, we had to select the {\it fully-connected} topology for the {\it Greedy} heuristics and for {\it mpShims} because the list of items to embark and the aircraft torque must be accessible for reading and writing by all child processes. This decision require an effective race condition management.

For the {\it mp3Dpacking} we used {\it parallel independent runs}, because there is no memory sharing.

@@

@@

In the first and third phases, we considered each pallet as a parallel process, taking advantage of nowadays multi-core computers, being executed in a multiprocessing run-time, accessing concurrently, reading and writing the list of items and the current value of the CG deviation.

@@

The main logic of the 3D packing procedure ($3Dpacker$) is based on an approximate algorithm like follows:

\begin{itemize}
	\item[1] From a list of items assigned to a pallet, items are sorted from the biggest to the smallest and placed in such an order on a pallet. Each item has 1 to 6 orientations to choose from in the moment of placement. The orientation procedure can select the best direction type among possible orientations.
	\item[2] A pivot point is used to determine item's position. The pivot is an (x, y, z) coordinate which represents a point in the pallet at which an attempt to pack an item will be made.
	\item[3] The back lower left corner of the item will be placed at the pivot. If the item cannot be packed at the pivot position then it is rotated until it can be packed at the pivot point or until we have tried all 6 possible rotation types (first fit).
	\item[4] If, after rotating it, the item still cannot be packed at the pivot point, then we move on to packing another item and add the unpacked item to a list of items that will be packed after an attempt to pack the remaining items is made.
	\item[5] The first pivot in the empty pallet is always (0, 0, 0). When one item can be placed into multiple optimal pivot point, the placement selection module can help make a choice (best fit).
\end{itemize}

@@

Let $p_i.Dh,\ p_i.Wh,\ p_i.H$ be parameters indicating the depth, width, and height of the pallet. We included letter $h$ at the end of parameters $p_i.Dh$ and $p_i.Wh$ to make them different from $p_i.D$ (distance from CG) and $p_i.W$ (weight limit).

@@

\begin{equation} \label{eq:3dpacking1}
	\mbox{minimize} \Big [ p_i.Dh \times p_i.Wh \times p_i.H - \sum_{j \gets 1}^{n^{\ i}} X_{ij} \times A_{j} \times t_j.d \times t_j.wh \times t_j.h \Big ]
\end{equation}

\begin{equation} \label{eq:3dpacking2}
	\sum_{j \gets 1}^{n^{\ i}} A_{j} \times t_j.d  \leq X_{ij} \times p_i.Dh\ i \in \{1,\ 2,\ 3,\ ...,\ m\}
\end{equation}

@@

\begin{equation} \label{eq:3dpacking5}
t_j.o^1 + t_j.o^2 + t_j.o^3 + t_j.o^4 + t_j.o^5 + t_j.o^6 \leq 1
\end{equation}

Equation \ref{eq:3dpacking1} minimizes the volume slack; equations \ref{eq:3dpacking2}, \ref{eq:3dpacking3}, and \ref{eq:3dpacking4} constrains the packing to the three dimensions; and \ref{eq:3dpacking5} guarantees that at most one orientation is chosen for the item.

@@

\begin{algorithm}[H]
	\caption{Multiprocessing 3D packing procedure}  \label{alg:3dpacking}

@@

\subsection{Minimization of the unloading costs}

@@

The instruction referred in Algorithm \ref{alg:tour}, line \ref{tour:minRampDist} implements this model.

@@

\subsection{Results before 3D packing}

@@

\begin{table}[H]
	\centering
	\caption{Volume surpluses, methods and scenarios results {\bf before} 3D packing}  \label{tab:results}
	\footnotesize
	%\scriptsize
	\begin{tabular}{crcccccccc}
		\toprule
		{\bf Volume}     &{\bf Resolution}&                              &\multicolumn{6}{c}{\bf Scenarios}                                                   &{\bf Normalized}\\
		\cmidrule{4-9}		
		{\it surplus}    &{\it method}    & {\bf Results}                &{\bf 1}    &{\bf 2}    &{\bf 3}    &{\bf 4}    &{\bf 5}    &{\bf 6}    &{\bf Speedup} \\
		\toprule	
		\multirow{6}{*}{$1.2$}&$Relaxed$  & $f$                &{\it 12.23}&{\it 8.67} &{\it 12.67}&{\it 15.36}&{\it 16.29}&{\it 16.20}&  \specialcell{{\bf 1.00}} \\%  8142
		&                      $MIP$      & {\it run time (s)} &     3     &     13    &     20    &     78   &      198  &     1103  &  1.00 \\% 439							
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $Shims$}        & $f$                &     10.34 &     8.50  &     12.42 &     15.06 &     15.97 &     15.88 &  0.96 \\% 7817
		&                                 & {\it run time (s)} &     < 1   &     3     &     5     &     20    &     58    &     352   &  1.00 \\% 439			
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $mpShims$}      & $f$                &   12.96   &   7.53    &  12.28    &   13.20   &  13.71    &   13.44   &  0.90\\% 7278
		&                                 & {\it run time (s)} &     < 1   &     2     &     2     &     8    &     22    &     120   &  {\bf 2.80}\\%
		\midrule		
		\multirow{6}{*}{$1.5$}&$Relaxed$  & $f$                &{\it 17.01}&{\it 12.45}&{\it 18.02}&{\it 22.03}&{\it 23.09}&{\it 22.35}&  \specialcell{{\bf 1.00}} \\%11495  
		&                       $MIP$     & {\it run time (s)} &     6     &     25    &     41    &     158   &      401  &     2161  &  1.00\\%					
        \cmidrule{2-10}		                       
        &\multirow{2}{*}{ $Shims$}        & $f$                &     14.68 &     12.21 &     17.67 &  21.01    &  22.64    &   21.91   &  0.96  \\%  11012
        &                                 & {\it run time (s)} &     < 1   &     2     &     3     &     12    &     33    &   209     &  1.00 \\% 260
        \cmidrule{2-10}		                      
        &\multirow{2}{*}{ $mpShims$}      & $f$                &    16.12  &     11.81 &     17.48 &  17.66    &   19.81   &  17.61    &  0.87 \\% 9966
        &                                 & {\it run time (s)} &     < 1   &     1     &     2     &     7     &   20      &  120      &  {\bf 1.70}\\% 151
		\midrule
		\multirow{6}{*}{$2.0$}&$Relaxed$  & $f$                &{\it 24.56}&{\it 17.47}&{\it 25.19}&{\it 27.94}&{\it 29.90}&{\it 27.60 }&  \specialcell{{\bf 1.00}} \\%15266  
		&                       $MIP$     & {\it run time (s)} &     6     &     25    &     41    &     158   &      401  &     2161  & 1.00\\%							
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $Shims$}        & $f$                &    21.31  &  17.35    &   24.24   &   27.39    &  29.31     &  27.06     &   0.96  \\%  14666
		&                                 & {\it run time (s)} &     < 1   &     2     &     4     &     14     &     35     &   220     &  1.00 \\% 260
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $mpShims$}      & $f$                &    24.33  &   17.26   &   24.70   &  24.81     &   27.71    &  24.93        &  0.94  \\% 14374
		&                                 & {\it run time (s)} &     < 1   &     2     &     3     &     9      &   26       &  151      &  {\bf 1.70}\\% 151
		\bottomrule	
	\end{tabular}
	\normalsize
\end{table}

@@

@@

We confirmed the hypothesis raised in \ref{hypotheses} and in line \ref{mpShims:minCGdev} of Algorithm \ref{alg:mpShims} that, in sequential mode, where one pallet is built at a time, torque constraints could be more restrictive than volume or weight constraints, which could stop the palletization process earlier, yielding a solution that may be not as good as it could be. This integer programming procedure inclusion noticeably improved all {\it Shims} results.

@@

\subsection{Results after 3D packing}

\vspace{2.0mm}
\begin{table}[H]
	\centering
\color{blue}{	
	\caption{Volume surpluses, methods and scenarios results {\bf after} 3D packing}  \label{tab:results2}
}
	\footnotesize
	\begin{tabular}{crcccccccc}
		\toprule
		{\bf Volume}     &{\bf Resolution}&                    &\multicolumn{6}{c}{\bf Scenarios}                                                   &{\bf Normalized}\\
		\cmidrule{4-9}		
		{\it surplus}    &{\it method}    & {\bf Results}      &{\bf 1}    &{\bf 2}    &{\bf 3}    &{\bf 4}    &{\bf 5}    &{\bf 6}    &{\bf Speedup} \\
		\toprule
		\multirow{6}{*}{$1.2$}&$Relaxed$  & $f$                &{\it 11.47}&{\it 7.24} &{\it 10.78}&{\it 12.06}&{\it 12.97}&{\it 12.44}&  \specialcell{{\bf 1.00}} \\%
		&                       $MIP$     & {\it run time (s)} &     6     &     25    &     41    &     158   &      401  &     2161  &  1.00\\%					
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $Shims$}        & $f$                &    9.05   &   5.11    &   9.75    &   10.96   &   12.86   &   13.25   &  0.96 \\% 7817
		&                                 & {\it run time (s)} &     16    &     68    &     94    &    323    &    719    &    3549   &  1.00 \\% 439			
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $mpShims$}      & $f$                &    12.14  &   7.29    &   10.36   &   12.63   &   14.04   &   13.92   &  0.90\\% 7278
		&                                 & {\it run time (s)} &     9     &     36    &     54    &     192   &     427   &    2118   &  {\bf 2.80}\\%		
		\midrule		
		\multirow{6}{*}{$1.5$}&$Relaxed$  & $f$                &{\it 15.83}&{\it 10.08}&{\it 15.14}&{\it 16.87}&{\it }&{\it }&  \specialcell{{\bf 1.00}} \\%11495  
		&                      $MIP$      & {\it run time (s)} &    10     &     43    &     72    &     262   &     58    &     352   &  1.00 \\% 439					
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $Shims$}        & $f$                &   11.55   &   7.61   &   12.28   &   15.00   &   17.55   &   17.57   &  0.96  \\%  11012
		&                                 & {\it run time (s)} &    25     &    107   &     156   &    498    &    1217   &   5743     &  1.00 \\% 260
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $mpShims$}      & $f$                &   17.05   &   10.01   &   15.01  &   16.75   &   20.09   &      &  0.87 \\% 9966
		&                                 & {\it run time (s)} &    18     &    64     &     101  &    351    &   828     &  120      &  {\bf 1.70}\\% 151
		\midrule
		\multirow{6}{*}{$2.0$}&$Relaxed$  & $f$                &{\it 23.73}&{\it }&{\it }&{\it }&{\it }&{\it  }&  \specialcell{{\bf 1.00}} \\%15266  
		&                       $MIP$     & {\it run time (s)} &    14     &     3     &     5     &     20    &     58    &     352   &  1.00 \\% 439					
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $Shims$}        & $f$                &   16.28   &      &      &       &       &       &   0.96  \\%  14666
		&                                 & {\it run time (s)} &    37     &     2     &     4     &     14     &     35     &   220     &  1.00 \\% 260
		\cmidrule{2-10}		                       
		&\multirow{2}{*}{ $mpShims$}      & $f$                &   24.52   &      &      &       &       &          &  0.94  \\% 14374
		&                                 & {\it run time (s)} &    27     &     2     &     3     &     9      &   26       &  151      &  {\bf 1.70}\\% 151
		\bottomrule	
	\end{tabular}
	\normalsize
\end{table}

\subsection{Results on the minimization of the unloading costs}
\label{improvement}
